---
title:  |
        | Tools for tidy analysis 
author: |
        | Christopher Skovron
        | Northwestern University 
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
    revealjs::revealjs_presentation:
          transition: none
          fig_height: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


## Tidying up the analysis stage



## Tools 

- `broom` - models
- `infer` - hypothesis testing
- `corrr` - tidy analysis of correlations 

## Additional modeling packages emerging 





## Tidy correlations with `corrr`

```{r}
#install.package('corrr')
library(corrr)
d <- correlate(mtcars)
d
```




## Using `infer` for tidy hypothesis testing

[`infer`](https://infer.netlify.com/reference/index.html) provides 





Set up mtcars for the example analysis - mutate vars to factors

```{r}
mtcars <- as.data.frame(mtcars) %>%
  mutate(cyl = factor(cyl),
          vs = factor(vs),
          am = factor(am),
          gear = factor(gear),
          carb = factor(carb))
```

## Difference in proportions: the old way




## Difference in proportions with `infer`

```{r}
mtcars %>%
  specify(am ~ vs, success = "1") %>%
  hypothesize(null = "independence") %>%
  generate(reps = 100, type = "permute") %>%
  calculate(stat = "diff in props", order = c("1", "0"))
```


## `infer` functions

`specify()` - 
`hypothesize()` - 
`generate()` - 
`calculate()` - 


## Difference in means: the old way



## Difference in means with `infer`

```{r}
mtcars %>%
  specify(response = mpg, explanatory = am) %>%
  generate(reps = 100, type = "bootstrap") %>%
  calculate(stat = "diff in means", order = c("1", "0"))
```










## Tidy bootstrapping with `broom`



## Bootstrapping the old way


## Bootstrapping with `broom`


## `broom` tidies up regression output 

- output from `lm`, `glm`, `t.test` is ugly and not standardized
- `broom` transforms into easy-to-use matrices
- makes manipulating model output much easier
- especially useful if you're simulating a lot of models 


## Example: weight and mileage in `mtcars`

```{r, echo=TRUE}
library(ggplot2)
data(mtcars)
ggplot(mtcars, aes(mpg, wt)) + geom_point()
```


## Fit an nls model

```{r, echo=TRUE}
nlsfit <- nls(mpg ~ k / wt + b, mtcars, start=list(k=1, b=0))
summary(nlsfit)
```

## model 
```{r}
ggplot(mtcars, aes(wt, mpg)) + geom_point() + geom_line(aes(y=predict(nlsfit)))
```


## We want to bootstrap the confidence intervals


```{r, echo=TRUE, message = FALSE}
library(dplyr)
library(broom)
set.seed(2014)
bootnls <- mtcars %>% bootstrap(100) %>%
    do(tidy(nls(mpg ~ k / wt + b, ., start=list(k=1, b=0))))
```

## We get a clean `data.frame` that summarizes each replication 

```{r, echo=TRUE}
bootnls
```

## use the percentile method to calculate uncertainty 

```{r, echo=TRUE}
alpha = .05
bootnls %>% group_by(term) %>% 
  summarize(low=quantile(estimate, alpha / 2),
  high=quantile(estimate, 1 - alpha / 2))
```

## Or you can use `augment` to visualize the uncertainty in the curve:

```{r, echo=TRUE}
bootnls_aug <- mtcars %>% bootstrap(100) %>%
    do(augment(nls(mpg ~ k / wt + b, ., start=list(k=1, b=0)), .))

ggplot(bootnls_aug, aes(wt, mpg)) + geom_point() +
    geom_line(aes(y=.fitted, group=replicate), alpha=.2)
```


## `broom` and "the secret weapon"

- Gelman's "secret weapon" 
- Fit the same model repeatedly to different datasets 

## Use `dotwhisker` and `broom` to use the secret weapon in a tidy way 




